---
title: "Homework 6"
author: "mm5951"
date: "`r Sys.Date()`"
output: github_document
---

````{r, include = FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(rvest)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

Solutions provided by teaching team.


## Problem 2

### Data wrangling & description 

First, I import the dataset using `read_csv()` and rename empty observations to "na" ("", "Unknown"). To describe the raw data, I use the `skimr::skim()` function.

```{r, warning = FALSE, message = FALSE}
homicide_df_raw = read_csv("./data/homicide-data.csv", na = c("","Unknown"))

skimr::skim(homicide_df_raw)
```

The "homicide_df_raw" dataset contains `r ncol(homicide_df_raw)` variables and `r nrow(homicide_df_raw)` observations. Details on the nature of its variables and summary values are found in the outputs above.

Then, I wrangle data as per problem instructions. This includes:

* Create a new "city_state" variable (e.g. “Baltimore, MD”) with `mutate()` and order by alphabetical orden using `str_c()`.
* Create a new "resolution" variable using the `case_when()` syntaxis, indicating whehter a case is resolved or not (those for which the disposition is “Closed without arrest” or “Open/No arrest”).
* Note one entry "Tulsa, AL" is excluded using `filter()`, as it is unclear whether it refers to Tulsa, Oklahoma or Birmingham, Alabama (this entry becomes apparent on the next section, under the "US_summary" data frame, and is then retroactively amended).

```{r}
homicide_df = homicide_df_raw %>% 
  mutate(city_state = str_c(city, state, sep = ", "),
         resolution = case_when(
           disposition == "Closed without arrest" ~ "unsolved",
           disposition == "Open/No arrest" ~ "unsolved",
           disposition == "Closed by arrest" ~ "solved"
         )) %>% 
  relocate(city_state) %>% 
  filter(city_state != "Tulsa, AL")
```


 

## Problem 3

### Data wrangling & description 
