---
title: "Homework 6"
author: "mm5951"
date: "`r Sys.Date()`"
output: github_document
---

````{r, include = FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(modelr)

library(rvest)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

Solutions provided by teaching team.


## Problem 2

### Data wrangling  

**PROMPT**: Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake. For this problem, limit your analysis those for whom victim_race is white or black. Be sure that victim_age is numeric.


First, I import the dataset using `read_csv()` and rename empty observations to "na" ("", "Unknown"). Then, I wrangle data as per problem instructions. This includes:

* Create a new "city_state" variable (e.g. “Baltimore, MD”) with `mutate()` and order by alphabetical orden using `str_c()`.
* Create a new "resolution" variable using the `case_when()` syntaxis, indicating whehter a case is resolved or not (those for which the disposition is “Closed without arrest” or “Open/No arrest”).
* Note one entry "Tulsa, AL" is excluded using `filter()`, as it is unclear whether it refers to Tulsa, Oklahoma or Birmingham, Alabama.
* **OTHER REQUESTS OF EXCLUSION & as.numeric**

```{r, warning = FALSE, message = FALSE}
homicide_df = read_csv("./data/homicide-data.csv", na = c("","Unknown")) %>% 
  mutate(city_state = str_c(city, state, sep = ", "),
         resolution = case_when(
           disposition == "Closed without arrest" ~ "unsolved",
           disposition == "Open/No arrest" ~ "unsolved",
           disposition == "Closed by arrest" ~ "solved"
         )) %>% 
  relocate(city_state) %>% 
  filter(city_state != "Tulsa, AL")
```

**CONCLUSION**: The "homicide_df_raw" dataset contains `r ncol(homicide_df_raw)` variables and `r nrow(homicide_df_raw)` observations. 

### Logistic regression 

**PROMPT**: For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. Save the output of glm as an R object; apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.

Now run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims. Do this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.


### Data visualization

**PROMPT**: Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.



## Problem 3

### Data wrangling  

First, I load using `read_csv()` and wrangle it so it's tidy for regression analysis.

Using `sum(is.na())` I check for missing data.

```{r}
birthweight = read_csv("./birthweight.csv") %>% 
  mutate(babysex = factor(if_else(babysex == 1, "male", "female")),
         frace = factor(recode(frace, '1' = "White", '2' = "Black", '3' = "Asian", 
                               '4' = "Puerto Rican", '8' = "Other", '9' = "Unknown")),
         mrace = factor(recode(mrace,'1' = "White", '2' = "Black", '3' = "Asian", 
                               '4' = "Puerto Rican", '8' = "Other", '9' = "Unknown")),
         malform = factor(recode(malform, '0' = "absent", '1' = "present")))

sum(is.na(birthweight))
```


### Linear regression model for birth weight

Next, I investigate a regression model for birth weight (`bwt`) against other potential associated variables. Since a baby's birth weight is typically with its length (`blenght`), I use a scatter plot as an exploratory analysis.

```{r}
birthweight %>% 
  ggplot(aes(x = blength, y = bwt)) + 
  geom_point(alpha = .5) +
  labs(title = "Scatter plot of baby's birth weight and baby's birth length") +
  theme(plot.title = element_text(hjust = 0.5))
```

As per the visualization, it appears that there might be a linear relationship between a baby's length and weight at birth. Thus, in a logsitic regression model "fit1" I will input `blength` as a predictor of the birth weight `bwt`. To assess its fitting, I will plot the model residuals against fitted values.

```{r}
fit1 = lm(bwt ~ blength, data = birthweight)
fit1 %>% 
  broom::tidy() %>% 
  knitr::kable()
birthweight %>% 
  modelr::add_residuals(fit1) %>% 
  modelr::add_predictions(fit1) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = .3) +
  labs(title = "Plot of model residuals against fitted values for fit1 model",
       x = "Fitted Values",
       y = "Residuals") +
  theme(plot.title = element_text(hjust = 0.5))
```

### Two alternative linear regression models for birth weight

In order to assess the fitting of the fit1 model, I will compare it to two alternative models, entailing:

* fit2: Using length at birth and gestational age as predictors (main effects only);
* fit3: Using head circumference, length, sex, and all interactions (including the three-way interaction) between these as predictors.

To do so, I use again the function `lm()` including the abovementioned variables in its arguments. Again, the model of residuals is ploted for both fit2 and fit3.

```{r}
fit2 = lm(bwt ~ blength + gaweeks,birthweight)

fit2 %>% 
  broom::tidy() %>% 
  knitr::kable()
birthweight %>% 
  modelr::add_residuals(fit2) %>% 
  modelr::add_predictions(fit2) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = .3) +
  labs(title = "Plot of model residuals against fitted values for fit2 model",
       x = "Fitted Values",
       y = "Residuals") +
  theme(plot.title = element_text(hjust = 0.5))

fit3 = lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, birthweight)

fit3 %>% 
  broom::tidy() %>% 
  knitr::kable()
birthweight %>% 
  modelr::add_residuals(fit3) %>% 
  modelr::add_predictions(fit3) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = .3) +
  labs(title = "Plot of model residuals against fitted values for fit3 model",
       x = "Fitted Values",
       y = "Residuals") +
  theme(plot.title = element_text(hjust = 0.5))
```

### Comparison of the fit1-3 linear regression models for birth weight




*That's it for homework 6!*

